---
title: Calculation of conversion factors for model parameters
author: Sasha D. Hafner
output: pdf_document
classoption: landscape
date: May 2021
---

```{r}
library(biogas)
```

Last run:

```{r}
Sys.time()
```


# Note on inputs
Chemical formulas from biogas package.
Document values of macromolecular composition.
These match the SoftwareX paper.

```{r}
biogas:::std.forms
```

Composition of manure based on macromolecular groups from Appendix S3 in the paper, as g/g VS.


# 1. COD:VS ratios
## A. Pig, degradable, no VFA

```{r}
pred <- predBg(mcomp = c(protein = 0.18, lipid = 0.12, carbohydrate = 0.39), 
               mass = 1, value = 'all')
pred
1/pred$COD
```

## C. Cattle, degradable, no VFA

```{r}
pred <- predBg(mcomp = c(protein = 0.15, lipid = 0.07, carbohydrate = 0.43), 
               mass = 1, value = 'all')
pred
1/pred$COD
```

# 2. Chemical formulas and CO2:COD ratios
These all use degradable VS plus VFA to give CO2 mass produced per g COD converted.
We need to reduce CO2 (and any other "energy reaction" products) by the fraction of substrate used for microbial biomass synthesis.
This is a bit sketchy because this biomass contributes to slurry OM (VS and COD).
But we only explicitly include it in the model for methanogens. 
So for the others (fermentation, aerobic, sulfate reduction?) we should include an effective fraction that represents slowly degradable or nondegradable biomass.

Yields in g COD/g COD.
Below I took about 20% off the yields from Table 3.1 in Rittmann and McCarty.

```{r}
yaer <- 0.1
yanaer <- 0.03
ysr <- 0.01
```

## A. Pig, degradable, with VFA

```{r}
pred <- predBg(mcomp = c(acetic = 0.08, protein = 0.18, lipid = 0.12, carbohydrate = 0.39), 
               mass = 1, value = 'all')
pred
1/pred$COD
```

Round for the following formula for degradable pig slurry organic matter (OM).

```{r}
fpd <- 'C16 H27 O8.7 N'
```
Anaerobic CO2.

```{r}
predBg(fpd, fs = yanaer, value = 'reactionc')
pred <- predBg(fpd, fs = 0.10, value = 'all')
pred
pred$mCO2 / pred$COD
```

So we have 0.49 CO2 per g COD.

Aerobic CO2.
Here we need the mass of CO2 assuming all C goes to CO2, per g COD.

```{r}
(1 - yaer) * molMass('CO2') * 16 / (calcCOD(fpd) * molMass(fpd))
```

Sulfate reduction:

```{r}
(1 - ysr) * molMass('CO2') * 16 / (calcCOD(fpd) * molMass(fpd))
```

## B. Cattle, degradable, with VFA

```{r}
pred <- predBg(mcomp = c(acetic = 0.04, protein = 0.15, lipid = 0.07, carbohydrate = 0.43), 
               mass = 1, value = 'all')
pred
1/pred$COD
```

Round for the following formula for degradable pig slurry organic matter (OM).

```{r}
fcd <- 'C16.6 H27.9 O10 N'
```
Anaerobic CO2.

```{r}
predBg(fcd, fs = yanaer, value = 'reactionc')
pred <- predBg(fcd, fs = yanaer, value = 'all')
pred
pred$mCO2 / pred$COD
```

0.57 g CO2 per g COD.

Aerobic CO2:

```{r}
(1 - yaer) * molMass('CO2') * 16.6 / (calcCOD(fcd) * molMass(fcd))
```

```{r}
(1 - ysr) * molMass('CO2') * 16.6 / (calcCOD(fcd) * molMass(fcd))
```


