---
title: "Oxygen mass transfer sensitivity in ABM"
author: "Sasha D. Hafner"
date: "`r Sys.Date()`"
fig_width: 6 
fig_height: 4 
output: pdf_document
---

```{r,echo=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4)
options(width = 85)
```

```{r,eval= TRUE}
library(ABM)
```

Default simulation

```{r}
out1 <- abm()
```
Total cumulative emission (g) and emission rate (g/d) are plotted below.

Completely shut off respiration, and also increase mass transfer coefficient by a factor of 10.

```{r}
out2 <- abm(add_pars = list(kl.oxygen = 0))
```

```{r}
out3 <- abm(add_pars = list(kl.oxygen = 5))
```

```{r}
plot(CH4_emis_rate ~ time, data = out1, type = 'l', ylab = 'Cumulative emission (g)', lwd = 2)
lines(CH4_emis_rate ~ time, data = out2, col = 'red', lty = 2)
lines(CH4_emis_rate ~ time, data = out3, col = 'orange', lty = 3)
```

```{r}
plot(CO2_emis_rate ~ time, data = out1, type = 'l', ylab = 'Cumulative emission (g)', lwd = 2)
lines(CO2_emis_rate ~ time, data = out2, col = 'red', lty = 2)
lines(CO2_emis_rate ~ time, data = out3, col = 'orange', lty = 3)
```

Conclude that respiration is minor sink for substrate, and uncertainty is not important.
Compare CH4/respiration COD consumption:

```{r}
plot(f_COD_CH4_cum ~ time, data = out3, type = 'l', col = 'gray45', lty = 2, ylab = 'Cumulative emission (g)', lwd = 2)
lines(f_COD_respir_cum ~ time, data = out1, lwd = 1)
```

```{r}
tail(out3[, c(1, which(grepl('^f_COD', names(out3))))])
```

Try increasing k_L 10x and decreasing the maximum depth.
To do that we will increase surface area from 11 m2 to 333 m2, for a maximum depth of 1 cm.

```{r}
out4 <- abm(add_pars = list(kl.oxygen = 5, area = 333))
```

```{r}
tail(out4[, c(1, which(grepl('^f_COD', names(out4))))])
```

Now respiration becomes more important.

```{r}
plot(CH4_emis_rate ~ time, data = out1, type = 'l', ylab = 'Cumulative emission (g)', lwd = 2)
lines(CH4_emis_rate ~ time, data = out4, col = 'orange', lty = 3)
```

```{r}
plot(CO2_emis_rate ~ time, data = out4, type = 'l', ylab = 'Cumulative emission (g)', col = 'blue')
lines(CO2_emis_rate ~ time, data = out1)
```

But even so, for default conditions CH4 emission is not affected.


